---
title: "initial_data_analysis"
author: "Freya Olsson"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
source('R/plotting_functions.R')

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figs")){dir.create("../Figs")}
```

## Read in data

```{r}
survey_results <- googlesheets4::read_sheet('https://docs.google.com/spreadsheets/d/1dKGzFRlMCs8bYsyo3gmbrZPAY4-bgWMH_L9T23iO6iE/edit?usp=drive_link', col_types = 'c') |> 
  filter(as_datetime(StartDate) >= as_datetime('2023-09-21')) 
  
questions <- googlesheets4::read_sheet('https://docs.google.com/spreadsheets/d/1dKGzFRlMCs8bYsyo3gmbrZPAY4-bgWMH_L9T23iO6iE/edit?usp=drive_link', col_types = 'c')[1,]
```

## Format data, add summary columns

```{r}
survey_results_formatted <- survey_results %>%
  mutate(Progress = as.numeric(Progress),
         Country_group = ifelse(Residence %in% c("North America", "Europe"),
                                Residence,
                                "Other"),
         Country_group = ifelse(is.na(Residence), NA, Country_group),
         Sector_group = ifelse(Sector == "Academia", "Academia", "Other"),
         Sector_group = ifelse(is.na(Sector), NA, Sector_group),
         Career = factor(Career, levels = c("Graduate student/early-career researcher", 
                                            "Mid-career faculty/research staff",
                                            "Late-career researcher/employee"))) %>%
  filter(Progress > 5)
```

## Bar graph of Experience with Data Analysis, facetted by career stage

```{r}
jpeg("Figs/Experience_by_career.jpeg", width = 7, height = 5, units = "in", res = 300)
plot_facet_result(column_id = "Experience", group_id = "Career", 
                  df = survey_results_formatted %>% 
                    mutate(Experience = str_wrap(Experience, 30)), 
                  questions = questions, as_percent = F, switch_x_y = T)
dev.off()
```

## Responses to question "As a data user, what kind of open data challenges have you encountered?"

-\> maybe \# standardised to percentages, and again, could be nice to visualise career stage in there (maybe with colours)?

```{r}
user_challenges <- sub("As a data user, what kind of open data challenges have you encountered. - ", "", questions[27:32])
names(user_challenges) <- colnames(survey_results_formatted)[27:32]

survey_results_formatted_challenge <- survey_results_formatted %>% 
  filter(`Role type` %in% c("Data user: people who analyze or use data that others collect and share (e.g., modeller)", "Both")) %>%
  pivot_longer(Challenges_1...27:Challenges_6...32, names_to = "User_challenge", values_to = "Frequency") %>%
  group_by(IPAddress) %>%
  filter(!sum(is.na(Frequency)) == 6) %>%
  ungroup() %>%
  mutate(User_challenge = user_challenges[User_challenge],
         Frequency = ifelse(is.na(Frequency), "NA", Frequency),
         Frequency = ifelse(Frequency == "Rare challenge", "Occasional challenge", Frequency),
         Frequency = factor(Frequency, levels = c("NA", "Never encountered","Occasional challenge", "Frequent challenge")) )
  
jpeg("Figs/Data_user_challenges.jpeg", width = 10, height = 5, units = "in", res = 300)
survey_results_formatted_challenge %>%
  mutate(User_challenge = str_wrap(User_challenge, 30)) %>%
  filter(!Career == "Other") %>%
  ggplot(aes(y = User_challenge, fill = Frequency)) +
  geom_bar(position = "fill") +
  facet_wrap(~Career) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  xlab("fraction of respondents") +
  ggtitle("Data user challenges") +
  scale_fill_manual(values = c("grey", viridis::viridis(4)), guide = guide_legend(reverse = TRUE))
dev.off()

user_unfacet <- survey_results_formatted_challenge %>%
  mutate(User_challenge = str_wrap(User_challenge, 30)) %>%
  filter(!Career == "Other") %>%
  ggplot(aes(y = User_challenge, fill = Frequency)) +
  geom_bar(position = "fill") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  xlab("Fraction of respondents") +
  ggtitle("Data user challenges") +
  scale_fill_manual(values = c("grey", viridis::viridis(3)), guide = guide_legend(reverse = TRUE))
```

## Bar graph of question "As a data user, what do you think are the minimum requirements for data providers to make their data and workflow sufficiently accessible?"

```{r}
survey_results_formatted_requirements <- survey_results_formatted %>%
  filter(`Role type` %in% c("Data user: people who analyze or use data that others collect and share (e.g., modeller)", "Both")) %>%
  mutate(Detailed_metadata = grepl("Detailed & high quality metadata", Requirements),
         Accessible = grepl("Accessible data file format and file size", Requirements),
         No_account = grepl("Directly accessible data, without requesting account", Requirements),
         Workflow = grepl("Publish reproducible workflow", Requirements),
         Available = grepl("Not necessary, available data is enough", Requirements)) %>%
  pivot_longer(Detailed_metadata:Available, names_to = "Requirement", values_to = "Response")

reqs <- c("Detailed_metadata", "Accessible", "No_account", "Workflow", "Available")
names(reqs) <- c("Detailed & high quality metadata", 
                 "Accessible data file format and file size", 
                 "Directly accessible data, without requesting account", 
                 "Publish reproducible workflow", 
                 "Not necessary, available data is enough")

jpeg("Figs/Data_user_requirements.jpeg", width = 7, height = 5, units = "in", res = 300)
survey_results_formatted_requirements %>%
  ggplot(aes(x = Requirement, fill = Response)) +
  geom_bar(position = "fill") +
  scale_x_discrete(breaks = reqs, labels = str_wrap(names(reqs), 20)) +
  theme(axis.text.x = element_text(angle = 90))
dev.off()
```

## Responses to question "As a data provider, what kind of open data challenges have you encountered?"

-\> maybe \# standardised to percentages?

```{r}
provider_challenges <- sub("As a data provider, what kind of open data challenges have you encountered. - ", "", questions[43:48])
names(provider_challenges) <- colnames(survey_results_formatted)[43:48]

survey_results_formatted_challenge2 <- survey_results_formatted %>% 
  filter(`Role type` %in% c("Data generator/provider: people who generate data through field/lab work or who collect and process large amounts of raw data to produce derived data products (e.g., remote sensing products)", "Both")) %>%
  pivot_longer(Challenges_1...43:Challenges_6...48, names_to = "Provider_challenge", values_to = "Frequency") %>%
  group_by(IPAddress) %>%
  filter(!sum(is.na(Frequency)) == 6) %>%
  ungroup() %>%
  mutate(Provider_challenge = provider_challenges[Provider_challenge],
         Frequency = ifelse(is.na(Frequency), "NA", Frequency),
         Frequency = factor(Frequency, levels = c("NA", "Never encountered","Occasional challenge", "Frequent challenge"))
         )
  
jpeg("Figs/Data_provider_challenges.jpeg", width = 10, height = 5, units = "in", res = 300)
survey_results_formatted_challenge2 %>%
  mutate(Provider_challenge = str_wrap(Provider_challenge, 40)) %>%
  filter(!Career == "Other") %>%
  ggplot(aes(y = Provider_challenge, fill = Frequency)) +
  geom_bar(position = "fill") +
  facet_wrap(~Career) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  xlab("Fraction of respondents") +
  ggtitle("Data provider challenges") +
  scale_fill_manual(values = c("grey", viridis::viridis(3)), guide = guide_legend(reverse = TRUE))
dev.off()

provider_unfacet <- survey_results_formatted_challenge2 %>%
  mutate(Provider_challenge = str_wrap(Provider_challenge, 40)) %>%
  filter(!Career == "Other") %>%
  ggplot(aes(y = Provider_challenge, fill = Frequency)) +
  geom_bar(position = "fill") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  xlab("Fraction of respondents") +
  ggtitle("Data provider challenges") +
  scale_fill_manual(values = c("grey", viridis::viridis(3)), guide = guide_legend(reverse = TRUE))

library(ggpubr)
jpeg("Figs/All_challenges.jpeg", width = 10, height = 5, units = "in", res = 300)
ggarrange(provider_unfacet, user_unfacet, labels = "auto", common.legend = T, legend = "bottom")
dev.off()
```

## Bar graphs for questions on Motivation and Entities

```{r}
# Motivation
survey_results_formatted_motivation <- survey_results_formatted %>%
  mutate(Advisor = grepl("Advisor requirement", Motivation),
         Personal = grepl("Personal desire to support open science", Motivation),
         Funding = grepl("Funding program requirement", Motivation),
         Journal = grepl("Journal requirement", Motivation),
         Other = grepl("Other", Motivation),
         Recognition = grepl("Recognition", Motivation)) %>%
  pivot_longer(Advisor:Recognition, names_to = "Motivation_type", values_to = "Response")

jpeg("Figs/Data_provider_motivation.jpeg", width = 7, height = 5, units = "in", res = 300)
survey_results_formatted_motivation %>%
  ggplot(aes(x = Motivation_type, fill = Response)) +
  geom_bar(position = "fill") +
  facet_wrap(~Career)
dev.off()

# Entities
#"Which of the following entities do you think has the greatest influence to promote publication of open data/workflows?"

jpeg("Figs/Open_data_entities.jpeg", width = 7, height = 5, units = "in", res = 300)
survey_results_formatted %>%
  ggplot(aes(x = Entities)) +
  geom_bar() +
  facet_wrap(~Career) +
  theme(axis.text.x = element_text(angle = 90))
dev.off()
```

```{r}
# Motivation
survey_results_formatted_metadata <- survey_results_formatted %>%
  mutate(Repo = grepl("data repository", Metadata),
         Funding = grepl("funding agency", Metadata),
         None = grepl("without", Metadata),
         Team = grepl("project team", Metadata)) %>%
  pivot_longer(Repo:Team, names_to = "Meta_type", values_to = "Response") %>%
  filter(!is.na(Response))

survey_results_formatted_metadata %>%
  ggplot(aes(x = Meta_type, fill = Response)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90))
```

