---
title: "Supplementary Text 2 - Survey distribution and results"
output: 
     pdf_document:
         extra_dependencies: ["booktabs"]
---

This supplemental document provides a list of networks that were contacted for survey distribution (Text 2.1) and a summary of survey results for each questions (Text 2.2).

# Text 2.1 Distribution

We distributed our survey via the following scientific networks:

-   Aquatic Ecosystem MOdeling Network - Junior (AEMON-J)

-   British Ecological Society (BES) Aquatic Ecology Special Interest Group

-   Coastal and Estuarine Research Federation (CERF)

-   Consortium of Universities for the Advancement of Hydrologic Sciences (CUAHSI)

-   Ecological Forecasting Initiative (EFI)

-   Ecolog listserv (Ecological Society of America; ESA)

-   German Society for Limnology (DGL)

-   International Society of Limnology (SIL)

-   Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE)

The survey was also distributed via social media as a post on Twitter/X and Bluesky.

# Text 2.2 Results Summary

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy=TRUE, echo = TRUE)
#Load packages
library(tidyverse)
library(lubridate)
library(htmltools)
source('R/table_functions.R')

#Load data
# Note: data are read in from an authenticated Google Sheet that is only available to approved individuals.}
google_drive_loc <- 'https://docs.google.com/spreadsheets/d/1dKGzFRlMCs8bYsyo3gmbrZPAY4-bgWMH_L9T23iO6iE/edit?usp=drive_link'

survey_results <- googlesheets4::read_sheet(google_drive_loc,
                                            col_types = 'c') |> 
  filter(as_datetime(StartDate) >= as_datetime('2023-09-21')) 
```

```{r format, include=FALSE}
#Format data
survey_results_formatted <- survey_results %>%
  mutate(Progress = as.numeric(Progress),
         Country_group = ifelse(Residence %in% c("North America",
                                                 "Europe"),
                                Residence,
                                "Other"),
         Country_group = ifelse(is.na(Residence),
                                NA,
                                Country_group),
         Sector_group = ifelse(Sector == "Academia",
                               "Academia",
                               "Other"),
         Sector_group = ifelse(is.na(Sector),
                               NA, Sector_group),
         Residence = ifelse(
           Residence ==
             'South America, Central America, and Caribbean',
           'S. and C. America',
           Residence)) %>%
  filter(Progress > 5)

  
questions <- googlesheets4::read_sheet(google_drive_loc, col_types = 'c')[1,]
```

## Summary of individual question responses

```{r tables, results='asis', warnings = F, message = F, echo = F}
columns <- survey_results_formatted[,18:55] %>% 
  select(!contains('_TEXT')) %>% 
  colnames()

checkbox <- c("Source", "Requirements", "Metadata", "Contacted", "Motivation")

for (col in columns) {
  if(col %in% checkbox){
    print(make_summary_table_checkbox(df = survey_results_formatted, column_id = col))
  } else {
    print(make_summary_table(df = survey_results_formatted, column_id = col))
  }
  
}
```
