---
title: "Supplementary Text"
output: 
     pdf_document:
         extra_dependencies: ["booktabs"]
---

This supplemental document provides a table of acronyms (Text S1), a list of networks that were contacted for survey distribution (Text S2.1) and a summary of survey results for each question (Text S2.2).

# Text S1: Acronyms

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy=TRUE, echo = TRUE)
#Load packages
library(tidyverse)
library(lubridate)
library(htmltools)
library(formatR)
library(kableExtra)
source('R/table_functions.R')

#Load data
# Note: data are read in from an authenticated Google Sheet that is only available to approved individuals.}
google_drive_loc <- 'https://docs.google.com/spreadsheets/d/1dKGzFRlMCs8bYsyo3gmbrZPAY4-bgWMH_L9T23iO6iE/edit?usp=drive_link'

survey_results <- googlesheets4::read_sheet(google_drive_loc,
                                            col_types = 'c') |> 
  filter(as_datetime(StartDate) >= as_datetime('2023-09-21')) 

table <- read_csv("Table to add to SI.csv", show_col_types = F)
```

\renewcommand{\arraystretch}{1.5}
```{r table, results='asis', warnings = F, message = F, echo = F}
caption <- "Acronyms used in this manuscript. References can be found in the reference list in the main manuscript."
knitr::kable(table, caption = caption, escape = F,
             linesep = "\\addlinespace")
```

\

\

# Text S2.1 Distribution

We distributed our survey via social media as a post on Twitter/X and Bluesky and via the following scientific networks:

-   Aquatic Ecosystem MOdeling Network - Junior (AEMON-J)

-   British Ecological Society (BES) Aquatic Ecology Special Interest Group

-   Coastal and Estuarine Research Federation (CERF)

-   Consortium of Aquatic Science Societies (CASS). CASS members include:

    -   American Fisheries Society

    -   Association for the Sciences of Limnology and Oceanography

    -   Coastal and Estuarine Research Federation

    -   Freshwater Mollusk Conservation Society

    -   International Association for Great Lakes Research

    -   North American Lake Management Society

    -   Phycological Society of America

    -   Society of Canadian Aquatic Sciences

    -   Society for Freshwater Science

    -   Society of Wetland Scientists

-   Consortium of Universities for the Advancement of Hydrologic Sciences (CUAHSI)

-   Ecological Forecasting Initiative (EFI)

-   Ecolog listserv (Ecological Society of America; ESA)

-   German Society for Limnology (DGL)

-   Global Lake Ecological Observatory Network (GLEON)

-   International Society of Limnology (SIL)

-   Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE)

# Text S2.2: Results Summary

```{r format, include=FALSE}
#Format data
survey_results_formatted <- survey_results %>%
  mutate(Progress = as.numeric(Progress),
         Country_group = ifelse(Residence %in% c("North America",
                                                 "Europe"),
                                Residence,
                                "Other"),
         Country_group = ifelse(is.na(Residence),
                                NA,
                                Country_group),
         Sector_group = ifelse(Sector == "Academia",
                               "Academia",
                               "Other"),
         Sector_group = ifelse(is.na(Sector),
                               NA, Sector_group),
         Residence = ifelse(
           Residence ==
             'South America, Central America, and Caribbean',
           'S. and C. America',
           Residence)) %>%
  filter(Progress > 5) %>%
  mutate(across(everything(), ~ sub("a,G", "a, G", .x)))

  
questions <- googlesheets4::read_sheet(google_drive_loc, col_types = 'c')[1,] %>%
  mutate(across(everything(), ~ sub("- Selected Choice", "", .x)))
```

Summary of individual question responses

\renewcommand{\arraystretch}{1}
```{r tables, results='asis', warnings = F, message = F, echo = F}
columns <- survey_results_formatted[,18:55] %>% 
  select(!contains('_TEXT')) %>% 
  colnames()

checkbox <- c("Source", "Requirements", "Metadata", "Contacted", "Motivation")

for (col in columns) {
  #Add explanatory text
  if(col == "Challenges_1...27"){
    cat('Note: "rare challenge" and "occasional challenge" from tables 7-12 below were combined into one category ("occasional challenge") in figure 2 of the main manuscript. This was done for consistency with the categories for data providers (tables 20-25).')
  }
  #Add manual line breaks
  if(col == "Ease_4" | col == "Challenges_6...48"){
    cat("&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  
&nbsp;  ")
  }
  #Print tables
  if(col %in% checkbox){
    print(make_summary_table_checkbox(df = survey_results_formatted, column_id = col))
  } else {
    print(make_summary_table(df = survey_results_formatted, column_id = col))
  }
  
}
```
